Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-06-15 articles/formalism/sections/discussion.tex
reordered text and made some changes to section on polarized filaments

These real space operators provide a spatially intuitive way of understanding the different decompositions of the Stokes vector on the sphere. We explicitly  demonstrated that all the real space operators have the common characteristic  of being separable into a azimuthal part that is band limit independent and radial part that incorporates all the band limit dependence. The azimuthal part of the operator is primarily responsible for the requisite spinorial decomposition, while the radial weights determine the non-local dependence of the construction of the resultant fields on the original fields. Using the self similarity property of the radial functions we define a non-locality parameter and argued that it is proportional to $\ell_{\rm max}^{-1}$. The scalar decomposition of the Stokes parameters being of primary interest,  we empirically derived the non-locality parameter $\beta_0$ for this operation and showed that the proportionality constant $\ell_{0}=22$ provided a good prediction for the angular distance at which the ${}_{\mm}f$ falls below 1\% of its maxima as seen in \fig{fig:rad_ker_decay}.

Our careful study of the real space operators reveals the dual interpretation of the operators as either the Green's function or a complex convolving beam depending on whether the kernels are expressed in terms of the forward rotation Euler angles or the inverse rotation Euler angles. While the convolution interpretation is a familiar one, the Green's function interpretation is a new way of looking at this operation. This new interpretation allows us to think of ${}_{+2}X$ as some spin-2 charge which radiates out a complex spin-0 scalar field $E+iB$. The resultant complex scalar maps can be then understood as arising from superposition of the radiating field emanating from all the spin charges on the sphere. The $E/B$ mode maps are merely the real and imaginary parts of this resultant complex scalar field.

>>>>

These real space operators provide a spatially intuitive way of understanding the different decompositions of the Stokes vector on the sphere. We explicitly  demonstrated that all the real space operators have the common characteristic  of being separable into a azimuthal part that is band limit independent and radial part that incorporates all the band limit dependence. The azimuthal part of the operator is primarily responsible for the requisite spinorial decomposition, while the radial weights determine the non-local dependence of the construction of the resultant fields on the original fields. Using the self similarity property of the radial functions we define a non-locality parameter and show that it is proportional to $\ell_{\rm max}^{-1}$. The scalar decomposition of the Stokes parameters being of primary interest,  we empirically derived the non-locality parameter $\beta_0$ for this operation and showed that the proportionality constant $\ell_{0}=22$ provided a good prediction for the angular distance at which the ${}_{\mm}f$ falls below 1\% of its maxima as seen in \fig{fig:rad_ker_decay}.

Our careful study of the real space operators reveals the dual interpretation of the operators as either the Green's function or a complex convolving beam depending on whether the kernels are expressed in terms of the forward or inverse rotation Euler angles. While the convolution interpretation is a familiar one, the Green's function interpretation is a new way of looking at this operation. This new interpretation allows us to think of ${}_{+2}X$ as some spin-2 charge which radiates out a complex spin-0 scalar field $E+iB$. The resultant complex scalar maps can be then understood as arising from superposition of the radiating field emanating from all the spin charges on the sphere. The $E/B$ mode maps are merely the real and imaginary parts of this resultant complex scalar field.

### 2018-06-15 articles/formalism/sections/kernel_paintbrush.tex
reordered text and made some changes to section on polarized filaments

The non-zero $B$ result is somewhat surprising given that the polarization pattern is symmetric to both horizontal and vertical reflections through the filament center.  However, unlike a circular ring, this filament is not a configuration with a definite parity.  Because the scalar polarization descriptions are coordinate independent, the $E/B$ patterns do not depend on the orientation of the filament.  A filament inclined at $45^\circ$ will have a similar $E/B$ pattern, but different reflection symmetries.

%\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but no $B$-mode signal is above the noise.  We predict that it should be there in higher fidelity data.}
\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but claim no $B$-mode signal. We predict that a B-mode signal from filaments should be present. A naive filament stacking analysis is only suited for making E-mode detections. Detecting B-modes from filaments requires a more careful edge stacked filaments analysis. While the detectability of B-mode signature from filaments from Planck data calls for a more careful assessment, it should be detectable in higher fidelity data.}

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern (note that this cancellation would not happen for an ellipse or any loop of non-contant radius of curvature).  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside.  The $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations. \revisit{The spiral filament showcased in \fig{fig:polfilaments} has a constant radius of curvature and thats the reason for not seeing a B-mode signal in intermediate regions of the spiral. For filaments that don't have a constant radius of curvature one expects to generate more B-mode power.} \comment{This is an important point right? So even if the magnetic field is perfectly aligned with the filament, one can generate B-mode signal in intermediate regions of the filament just by having the filament with changing radius of curvature. There is no reason for filaments to have a constant radius of curvature.}

\revisit{Note that we have made the simplifying assumption that the polarization headless vectors are perpendicular to the magnetic field which is always assumed parallel to the tangent to the filament at any given point and this results in generating dominantly an E-mode signal while the B-mode signal is present only near the edges. By similar arguments it can be shown that that polarization pattern will be dominated by B-modes if the polarization vector is aligned at $+45$ (or $-45$) degrees to the tangent to the filament while the E-mode signal would only be present at the edges. Therefore a careful study of the $E/B$ mode power in filaments can provide insights into the orientation of the magnetic field with respect to the length of the filament.}

>>>>

The non-zero $B$ result is somewhat surprising given that the polarization pattern is symmetric to both horizontal and vertical reflections through the filament center.  However, unlike a circular ring, this filament is not a configuration with a definite parity.  Since the scalar description of polarization is coordinate independent, the $E/B$ patterns do not depend on the orientation of the filament.  A filament inclined at $45^\circ$ will have a similar $E/B$ pattern, but different reflection symmetries.

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern. Here it is important to note that this cancellation will not happen for an ellipse or any loop of non-contant radius of curvature.  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside. While the $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations, note that it is non-vanishing in the intermediate regions due to the varying radius of curvature of the spiral \footnote{ \revisit{Why do we use this specific spiral - Its easy to compute the tangent to the spiral at any point.}}. \comment{If this is correct, then one expects to see a higher B-mode signal in intermediate regions of the spiral, as compared to the B-mode signal in the intermediate regions of the linear filament. This  then should be more visible in a logarithmically scaled version of Fig. 8.}

\revisit{Note that we have made the simplifying assumption that the polarization headless vectors are perpendicular to the magnetic field which is always assumed parallel to the tangent to the filament at any given point. This results in generating dominantly an E-mode signal while the B-mode signal is present only near the edges. By similar arguments it can be shown that that polarization pattern will be dominated by B-modes if the polarization vectors were aligned at $+45$ (or $-45$) degrees to the tangent to the filament while the E-mode signal would only be present at the edges. One way to get this configuration is to change the orientation of the magnetic field vector with respect to the tangent to the filament. Therefore a careful study of the $E/B$ mode power in filaments can provide insights into the orientation of the magnetic field with respect to the length of the filament.}

%\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but no $B$-mode signal is above the noise.  We predict that it should be there in higher fidelity data.}
\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but claim no $B$-mode signal. We predict that a B-mode signal from filaments should be present, since they have a finite length and a varying radius of curvature. A naive filament stacking analysis is only suited for making E-mode detections. Detecting B-modes from filament edges will require a more careful edge stacked filament analysis. While the detectability of B-mode signature from filaments from Planck data calls for a more careful assessment, it should be detectable in higher fidelity data.}

\revisit{In \cite{Zaldarriaga2001a} it is argued that linear filaments of infinite length in which the configuration of the polarization vector is parallel or perpendicular to the length of the filament cannot produce B-mode power. While this conclusion is true for filaments of infinite length, in reality we expect filaments that are observed on the sky to have a finite length and in this case one expects a definite B-mode pattern from the edges of the filament. One also does not expect that filaments have a constant radius of curvature and this will result in an additional generation of B-mode power. Therefore our study indicates results which subtly differ from those presented in \cite{Zaldarriaga2001a} and these differences can be attributed to the more realistic assumptions about the geometry of the filaments made in our analysis.} \comment{Move to conclusions?}
